Import the study area extent to get the bounding box

```{r}
library(sf)

study_area <- st_read("data/projected_utm37n/study_area_straightened.gpkg")
st_bbox(study_area)
```

The FABDEM data is downloaded as folders per 10x10 degrees, which contain files for each 1x1 degree square. I've already separated the 12 files we need (by looking at the data in qgis) and put them in separate folder

Read them in while also merging them:

```{r}
install.packages("terra")
library(tidyverse)
library(stringr)
library(rgdal)
library(terra)


mergeList <- function(rasList){
  #Internal function to make a list of raster objects from list of files.
  ListRasters <- function(list_names) {
    raster_list <- list() # initialise the list of rasters
    for (i in 1:(length(list_names))){ 
      grd_name <- list_names[i] # list_names contains all the names of the images in .grd format
      raster_file <- raster::raster(grd_name)
    }
    raster_list <- append(raster_list, raster_file) # update raster_list at each iteration
  }
  
  #convert every raster path to a raster object and create list of the results
  raster.list <-sapply(rasList, FUN = ListRasters)
  
  # edit settings of the raster list for use in do.call and mosaic
  names(raster.list) <- NULL
  #####This function deals with overlapping areas
  raster.list$fun <- sum
  
  #run do call to implement merge over the list of raster objects.
  mos <- do.call(raster::merge, raster.list)
  
  #set crs of output
  crs(mos) <- crs(x = raster(rasList[1]))
  
  return(mos)
}

#use the function on the raster files 
raster_files <- list.files(path ="data/dem's/FABDEM_S_nefud",pattern = ".tif$",full.names = TRUE )
fabdem <- mergeList(raster_files)

#crop to study area
e <- extent(study_area)
fabdem_cropped <- crop(fabdem, e )
plot(fabdem_cropped)
writeRaster(fabdem_cropped, "data/dem's/fabdem_cropped.tif", overwrite = TRUE)
```

Read in the cop30dem and map both layers

```{r}
cop_dem30 <- raster("data/dem's/rasters_COP30/output_COP30_15km.tif")
cop_dem30_5km <- raster("data/dem's/rasters_COP30/output_COP30_5km_buffer.tif")
library(tmap)
tmap_mode("view")
tm_shape(fabdem_cropped)+
  tm_raster(style = "cont", palette = "PuOr", legend.show = TRUE)+
  tm_shape(cop_dem30)+
  tm_raster(style = "cont", palette = "PuOr", legend.show = TRUE)+
  tm_scale_bar()

utm37n <- CRS("+init=epsg:32637")
#can't use projectRaster because it uses too much memory. Need to use gdalUtils gdalwarp - BUT- gdalUtils is not available for this version of R (need to use an older version)
cop_dem_proj <- projectRaster(cop_dem30, crs=utm37n, res=30)
study_area_buff <- st_buffer(study_area, 5000, endCapStyle = "SQUARE", joinStyle = "MITRE", mitreLimit = 3)
cop_dem_5km_proj <- projectRaster(cop_dem30_5km, crs=utm37n, res=30)
writeRaster(cop_dem_proj ,"data/projected_utm37n/cop_dem30_15km_utm37n.tiff")
cop_dem <- raster("data/projected_utm37n/cop_dem30_utm37n.tif")
plot(cop_dem)
st_write(study_area_buff, "data/projected_utm37n/study_area_straightened_5km.gpkg")
```

Decided to just use the COP DEM30 because there are not may buildings or vegetation and the FABDEM was flattening the slopes of jebels (used raster calculator and compared the difference map in qgis)

DEM derivatives calculation
TRI and TPI
```{r}
#read in the utm37n 5km buffer of dem
dem <- rast("data/projected_utm37n/copdem30_5km_straight.tif")

#tri<- terrain(x = dem, opt = c("slope","aspect",  "tpi"), neighbors = 4)

ter.par[tpi]

tpi <- terrain(x = dem, v = "TPI", neighbors = 8)
plot(tpi)
#use raster::terrain to calucltae tri and tpi
#queens case neighbourhood
f_3<- matrix(1, nrow=3, ncol=3)
f_5<- matrix(1, nrow=5, ncol=5)
f_15<- matrix(1, nrow=15, ncol=15)
x <- dem

#calculating it with focal so you can have different sized windows
f <- matrix(1, nrow=3, ncol=3)

TRI <- focal(x, w=f, fun=function(x, ...) sum(abs(x[-5]-x[5]))/8)
TPI <- focal(x, w=f, fun=function(x, ...) x[5] - mean(x[-5]))
rough <- focal(x, w=f, fun=function(x, ...) max(x) - min(x), na.rm=TRUE)

tr_3 <- focal(x, w=f_3, fun=function(x, ...) sum(abs(x[-5]-x[5]))/8)
tr_5 <- focal(x, w=f_5, fun=function(x, ...) sum(abs(x[-5]-x[5]))/8)
tr_15 <- focal(x, w=f_15, fun=function(x, ...) sum(abs(x[-5]-x[5]))/8)

#need to convert to spatraster - think it used raster::focal, not terra::focal
tr_3_spat <- rast(tr_3)
tr_5_spat <- rast(tr_5)
tr_15_spat <- rast(tr_15)
#stack them
tri_stack <- c(tr_3_spat, tr_5_spat, tr_15_spat)
#set their names
#maybe do this instead?
x <- rast(tri_stack)
names(tri_stack) <- c("tri_3", "tri_5", "tri_15")
#export the stack 
terra::writeRaster(tri_stack, "data/projected_utm37n/tri_stack.tif")

#read back in as a rasterbrick
y <- raster::brick("data/projected_utm37n/tri_stack.tif") 

#writeRaster(tri, "data/projected_utm37n/tri.tif")

tpi_3 <- terra::focal(x, w=f_3, fun=function(x, ...) x[5] - mean(x[-5]))
tpi_5 <- terra::focal(x, w=f_5, fun=function(x, ...) x[5] - mean(x[-5]))
tpi_15 <- terra::focal(x, w=f_15, fun=function(x, ...) x[5] - mean(x[-5]))

#stack them
tpi_3_spat <- rast(tpi_3)
tpi_5_spat <- rast(tpi_5)
tpi_15_spat <- rast(tpi_15)
tpi_stack <- c(tpi_3_spat, tpi_5_spat, tpi_15_spat)
#set their names

names(tpi_stack) <- c("tpi_3", "tpi_5", "tpi_15")
#export the stack 
terra::writeRaster(tpi_stack, "data/projected_utm37n/tpi_stack.tif")

#Terrain roughness
rough_3 <- terrain(x = dem, v= "roughness", neighbors = 8)
rough_5 <- terra::focal(x, w=f_5, fun=function(x, ...) max(x) - min(x), na.rm=TRUE)
rough_15 <- terra::focal(x, w=f_15, fun=function(x, ...) max(x) - min(x), na.rm=TRUE)

r_stack <- c(rough_3, rough_5, rough_15)
#set their names
#maybe do this instead?
x <- rast(r_stack)
names(r_stack) <- c("rough_3", "rough_5", "rough_15")
#export the stack 
terra::writeRaster(r_stack, "data/projected_utm37n/roughness_stack.tif")

y <- rast("data/projected_utm37n/roughness_stack.tif")
names(y)
plot(y$rough_3)
```
Rasterise Geo maps
```{r}
geo <- st_read("../../saudi_geological_maps/Geology_souther_nefud_final.sqlite")
geo1 <- geo$abbreviation
geo1<- geo[, c("abbreviation")]
geo_rast <- st_rasterize(geo1%>% dplyr::select(abbreviation, GEOMETRY))
plot(geo_rast)
```

