---
title: "mustatil_analysis"
author: "Amy Hatton"
date: "26/07/2022"
output: html_document
---

```{r}
library(sf)
library(tidyverse)
library(terra)
library(raster)
library(tmap)
library(tmaptools)
library(stars)
library(maptools)
library(spatstat)
```
# Read in data

```{r}
mustatils <- st_read("data/projected_utm37n/mustatils_with_size_metrics.gpkg")
dem <- raster("data/projected_utm37n/copdem30_5km_straight.tif")
#lakes <- st_read("data/water_paul_gpkg/psb_compiled_lake_extents_with_jubbah.gpkg")
#rivers <- st_read("data/water_paul_gpkg/hydroshed_100k_not_in_dunes.gpkg")
#deposits <- st_read("data/water_paul_gpkg/psb_modelled_deposits_with_extents.gpkg")
study_area <- st_read("data/projected_utm37n/study_area_straightened.gpkg")
geo <- st_transform(st_read("../../saudi_geological_maps/digitised_map_whole_study_area_final.shp"), crs=32637)
slope <- raster("data/projected_utm37n/slope.tif")
aspect <- raster("data/projected_utm37n/aspect.tif")
tpi <- raster("data/projected_utm37n/tpi.tif")
tri <- raster("data/projected_utm37n/tri.tif")
```

# Rose diagram for direction of mustatils
```{r}
#prep data
hist(mustatils$angle, plot=TRUE, bin=10)
summary(mustatils$angle)

ggplot() + 
  geom_line(aes(x = c(180, 0), y = c(4, 4)), colour = "gray75") +
  geom_line(aes(x = c(180, 0), y = c(8, 8)), colour = "gray75") +
  geom_line(aes(x = c(180, 0), y = c(12, 12)), colour = "gray75") +
  geom_line(aes(x = c(180, 0), y = c(16, 16)), colour = "gray75") +
  geom_line(aes(x = c(180, 0), y = c(20, 20)), colour = "gray75") +
  geom_vline(aes(xintercept = 0:6 * 30), colour = "gray75") +
  geom_histogram(data = mustatils, aes(x = angle), 
                 position = "stack", colour = "black", binwidth = 10,
                 boundary = 0) + 
  coord_polar(start = 3 * pi / 2, direction = -1) + 
  scale_x_continuous(limits = c(0, 360), breaks = 0:6 * 30) +
  scale_y_continuous(limits = c(0, 20)) +
  theme_bw() + 
  theme(panel.border = element_blank(), 
        legend.margin = margin(unit(c(0, 5.5, 100, 5.5), "pt")),
        axis.title.y = element_text(hjust = 0.75),
        axis.title.x = element_text(vjust = 5),
        plot.margin = margin(unit(c(50, 5.5, -100, 5.5), "pt")),
        panel.grid = element_blank()) +
  labs(title = "", y = "Count")

ggsave( "plots/rose_plot.pdf")
```


#Get Geology values for each mustatil

## First fix up geology attributes 

```{r}
names(geo)
#fix geology attributes
geo_df <- geo %>% dplyr::select(abbreviati, age, geology) %>% 
  rename(geology_abbr = abbreviati) %>% 
  tibble::rowid_to_column( "ID")

geo_attr <- read.csv("../../mustatil_paper/geology1_consolidated_paper.csv", sep = ";") 

#rename columns in geo_attr for joining
geo_attr <- geo_attr %>% 
  rename(geology_1 = Geology, age_1 = Age)

geo_attr <- geo_attr %>% 
  mutate(age_1 = tolower(age_1)) %>% 
  mutate(geology_1 = tolower(geology_1))

#geom <- geo %>% 
  #dplyr::select(geom)
geo_join <- dplyr::left_join(geo_df, geo_attr)

#fix nulls
geo_join <- geo_join %>% 
  mutate(age = coalesce(age, age_1),
         geology = coalesce(geology, geology_1)) 


#fix geology and age for Qt (not sure why buty there are a bunch of NA's for that geology)
geo_join <- geo_join %>% 
  mutate(geology_abbr = case_when(geology =="Devonian-Silurian-Ordovician undivided" ~ "DSO", TRUE ~ geology_abbr ))%>% 
  mutate(geology = case_when(geology_abbr == "Qt" ~ "Alluvial deposits", TRUE ~ geology)) %>% 
  mutate(age = case_when(geology_abbr == "Qt" ~ "quaternary", TRUE ~ age)) %>% 
  mutate(age = tolower(age)) %>% 
  mutate(geology = tolower(geology))

#issue with empty geometries
#fixed_geo <- geo_join %>% filter(!st_is_empty(.))

#check <- st_is_valid(geo_join)
#check <-as.data.frame(check)

#replace dash's with commmas
geo_join<- geo_join %>% 
  mutate(across('geology', str_replace, '-', ','))
#isolate all geology that is not quaternary
non_quart_geo <- filter(geo_join, age != "quaternary")
sandstone_geo <- geo_join %>% 
  filter(grepl("sandstone", geology))
```
Check the two geology maps to make sure they're right
```{r}

#all good
tmap_mode("view")

tm_shape(sandstone_geo)+
  tm_polygons("geology_abbr")

tm_shape(non_quart_geo)+
  tm_polygons("geology_abbr")

```

```{r, eval = FALSE}
library(magrittr)
#convert mustatil to points
m_points <- st_centroid(mustatils)

poly <- geo_join %>% 
  dplyr::select("geology_abbr") 
st_crs(poly)

#geology at the centre of each mustatil - make a barplot or some summary statistic of this
m_geology <- st_intersection(m_points, poly)

```

#Fix the data (only need to run this once)
```{r, eval=FALSE}
dem<- rast(dem)
x <- rast(dem)
res(x) <- 30
# make sure resolution is exactl 30mx30m
dem_res30 <- terra::resample(dem, x)
study_a_vect <- vect(study_area)
dem_cropped = crop(dem_res30, study_a_vect)
dem_final = mask(dem_cropped, study_a_vect, overwrite=TRUE)
writeRaster(dem_final, "data/cleaned_data/rast/dem.tif")
```

#Read in fixed data

```{r}
dem <- rast("data/cleaned_data/rast/dem.tif")
```

Rasterise geology and water maps
```{r, eval=FALSE}
#geology
non_quart_geo_rast <- rasterize(vect(non_quart_geo), dem, field=1)
sandstone_geo_rast <- rasterize(vect(sandstone_geo), dem, field=1)
writeRaster(non_quart_geo_rast, "data/cleaned_data/rast/non_quart_geo.tif")
writeRaster(sandstone_geo_rast, "data/cleaned_data/rast/sandstone_geo.tif")
#water maps
rivers_r <- rasterize(vect(rivers), dem, field=1)
lakes_r <- rasterize(vect(lakes), dem, field=1)
deposits_r <- rasterize(vect(deposits), dem, field=1)
#not working properly - just gives a raster with values of two export and do it in grass
#need to change NA's to zeros!
lakes_na <- terra::ifel(is.na(lakes_r), 0, lakes_r)
rivers_na  <- terra::ifel(is.na(rivers_r ), 0, rivers_r )
deposits_na<- terra::ifel(is.na(deposits_r ), 0,deposits_r )


#add rasters d
#still not working - did it in grass
#import again
lakes_rivers_r  <- rast("data/cleaned_data/rast/lakes_rivers.tif")
deposits_rivers_r  <- rast("data/cleaned_data/rast/deposits_rivers.tif")
#convert 0 back to na 
lakes_rivers_na  <- terra::ifel(lakes_rivers_r == 0, NA, lakes_rivers_r)
deposits_rivers_na <- terra::ifel(deposits_rivers_r == 0, NA, deposits_rivers_r )

```
#Create Distance maps (read in r keeps crashing)
```{r}
#d_rivers <- distance(rivers_r)
#read them in because r was crashing
dist_lakes <- raster("data/cleaned_data/rast/dist_lakes.tif")
dist_rivers <- raster("data/cleaned_data/rast/dist_rivers.tif")
dist_deposits <- raster("data/cleaned_data/rast/dist_deposits.tif")
dist_lakes_rivers <- raster("data/cleaned_data/rast/dist_lakes_rivers.tif")
dist_deposits_rivers <- raster("data/cleaned_data/rast/dist_deposits_rivers.tif")

#create distance maps for geology
#first convert to raster
sandstone_geo_rast <- terra::rasterize(sandstone_geo, dem)
dist_sandstone_geo <- terra::distance(sandstone_geo_rast)
non_quart_geo_rast <- terra::rasterize(non_quart_geo, dem)
dist_non_quart_geo <- terra::distance(non_quart_geo_rast)
writeRaster(dist_sandstone_geo, "data/cleaned_data/rast/dist_sandstone_geo.tif")
writeRaster(dist_non_quart_geo, "data/cleaned_data/rast/dist_non_quart_geo.tif")
```


#Create ppp mustatils
```{r}
m_points <- st_centroid(mustatils)
coords <- as.data.frame(st_coordinates(m_points))
m_owin <- as.owin(study_area)
sppp <- ppp(x=coords$X, y=coords$Y, window=m_owin)
```
Assessing point distribution first steps:
KDE
```{r}
#check bandwidth #366.9m

bw.diggle(sppp)
#another method nearest neighbour distance *3 = ~9000m
nnd<- nndist(sppp)
mean(nnd)
?density.ppp
mkde350 <- density.ppp(sppp, sigma=350)
mkde1000 <- density.ppp(sppp, sigma=1000)
mkde5000 <- density.ppp(sppp, sigma=5000)
mkde9000 <- density.ppp(sppp, sigma=9000)


#plot to pdf
pdf("plots/kde.pdf", width=12, height=3)
par(mfrow = c(1,3))
plot(mkde1000,  main="Kernel density surface of sites (sigma=1km)")
points(sppp, pch=1, cex=0.7, col = "grey80", lwd=1.3)

plot(mkde5000,  main="Kernel density surface of sites (sigma=5km)")
points(sppp, pch=1, cex=0.7, col = "grey80", lwd=1.3)

plot(mkde9000,  main="Kernel density surface of sites (sigma=9km)")
points(sppp, pch=1, cex=0.7, col = "grey80", lwd=1.3)

dev.off()
```

#First Order Model
##check out different variable

```{r}
#elevation
dem <- raster("data/cleaned_data/rast/dem.tif")
elev_im <- as.im.RasterLayer(dem)
elev.rh <- rhohat(sppp, elev, confidence=0.95)
#800 to 1400m
hist.im(elev_im)
pdf("plots/rhohat/elevation.pdf", width =7, height = 4)
plot(elev.rh, main ="",xlab="Elevation (m)",xlim=c(800,1400), legend=FALSE)

#lakes
lakes_im <- as.im.RasterLayer(dist_lakes)
lakes.rh <- rhohat(sppp, d_lakes_im, confidence=0.95)
#0 to 40km
hist.im(lakes_im)
pdf("plots/rhohat/lakes.pdf", width =7, height = 4)
plot(lakes.rh, main ="", xlab="Distance to lakes (m)",xlim=c(0,40000), legend=FALSE)

#rivers
rivers_im <- as.im.RasterLayer(dist_rivers)
rivers.rh <- rhohat(sppp, rivers_im, confidence=0.95)
#0 to 66km
hist.im(rivers_im)
pdf("plots/rhohat/rivers.pdf", width =7, height = 4)
plot(rivers.rh, main ="", xlab="Distance to rivers (m)",xlim=c(0,66000), legend=FALSE)

#deposits
deposits_im <- as.im.RasterLayer(dist_deposits)
deposits.rh <- rhohat(sppp, deposits_im, confidence=0.95)
#0 to 38km
hist.im(deposits_im)
pdf("plots/rhohat/deposits.pdf", width =7, height = 4)
plot(deposits.rh, main ="", xlab="Distance to deposits (m)",xlim=c(0,38000), legend=FALSE)

#rivers and lakes
lakes_r_im <- as.im.RasterLayer(dist_lakes_rivers)
lakes_r.rh <- rhohat(sppp, lakes_r_im, confidence=0.95)
#0 to 20km
hist.im(lakes_r_im)
pdf("plots/rhohat/lakes_rivers.pdf", width =7, height = 4)
plot(lakes_r.rh, main ="", xlab="Distance to lakes and rivers (m)",xlim=c(0,20000), legend=FALSE)

#rivers and deposits
deposits_r_im <- as.im.RasterLayer(dist_deposits_rivers)
deposits_r.rh <- rhohat(sppp, deposits_r_im , confidence=0.95)
#0 to 66km
hist.im(deposits_r_im)
pdf("plots/rhohat/deposits_r.pdf", width =7, height = 4)
plot(deposits_r.rh, main ="", xlab="Distance to deposits adn rivers (m)",xlim=c(0,66000), legend=FALSE)


#sandstone
sandstone_im <- as.im.RasterLayer(dist_sandstone_geo)
sandstone.rh <- rhohat(sppp, sandstone_im, confidence=0.95)
#0 to 50km
hist.im(sandstone_im)
pdf("plots/rhohat/sandstone.pdf", width =7, height = 4)
plot(sandstone.rh, main ="", xlab="Distance to sandstone (m)",xlim=c(0,50000), legend=FALSE)

#non quarternary geology
non_quart_im <- as.im.RasterLayer(dist_non_quart_geo)
non_quart.rh <- rhohat(sppp, non_quart_im, confidence=0.95)
#0 to 50km
hist.im(non_quart_im )
pdf("plots/rhohat/non_quart_2.pdf", width =7, height = 4)
plot(non_quart.rh, main ="", xlab="Distance to non quarternary geology (m)",xlim=c(0,30000), legend=FALSE)

#tpi
tpi_im <- as.im.RasterLayer(tpi)
tpi.rh <- rhohat(sppp, tpi_im, confidence=0.95)
#-8 to +10
hist.im(tpi_im )
pdf("plots/rhohat/tpi.pdf", width =7, height = 4)
plot(tpi.rh, main ="", xlab="TPI",xlim=c(-8,10), legend=FALSE)

#tri
tri_im <- as.im.RasterLayer(tri)
tri.rh <- rhohat(sppp, tri_im, confidence=0.95)
#0 to 16
hist.im(tri_im )
pdf("plots/rhohat/tri.pdf", width =7, height = 4)
plot(tri.rh, main ="", xlab="TRI",xlim=c(0,16), legend=FALSE)

#slope
slope_im <- as.im.RasterLayer(slope)
slope.rh <- rhohat(sppp, slope_im, confidence=0.95)
#0 to 40
hist.im(slope_im )
pdf("plots/rhohat/slope.pdf", width =7, height = 4)
plot(slope.rh, main ="", xlab="Slope",xlim=c(0,40), legend=FALSE)

#aspect
aspect_im <- as.im.RasterLayer(aspect)
aspect.rh <- rhohat(sppp, aspect_im, confidence=0.95)
#0 to 360
hist.im(aspect_im )
pdf("plots/rhohat/aspect.pdf", width =7, height = 4)
plot(aspect.rh, main ="", xlab="Aspect",xlim=c(0,360), legend=FALSE)
```



